Exploring the linear separablity of syntactic and semantic information in BERT embeddings

### 2022-06-09 paper/paper.tex
adding result

In \citep{disentangle} the authors use paraphrase pairs and new target syntax to train a semantic encoder, syntactic encoder and decoder to learn separate representations of the semantic and syntactic information contained in BART embeddings, in order to create semantically equivalent paraphrases with the new syntactic structure. Alongside the encoders they also train an adversarial syntax discriminator to try and predict the source syntax from the semantic embeddings, thus encouraging the disentanglement of the semantic and syntactic information by training the semantic embedder to remove as much syntactic information as possible. Their results show that one can achieve some removal of syntactic information from semantic embeddings, so disentanglement of some information is possible. Though they do not achieve perfect separation of the two information sets. Other non-linear approaches to syntactic-semantic information disentanglement have been carried out in \citep{multiDis}

@@

%Need to update this section to reflect the new dataset
We use the English gold standard of Parallel Meaning Bank v4.0 \citep{pmbData} to test the linear separability of the semantic and syntactic information in word-level BERT embeddings. This data set contains word-level semantic tags and NAME FEATURE USED FOR CCG DERIVATIONS %Fill this in
 that we parse using CITE CCG PARSER %Fill this in
 to develop word-level CCG tags, both of which are used to train the linear classifiers used for probing and evaluation. The gold-standard data set is comprised of XX words. %Fill this in
 %Do we drop anything or do any other pre-processing that I missed here?

@@

\begin{table*}[h]
    \centering
    \begin{tabular}{llll}
    \hline
    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantic Labeling} \\
    \hline
    %\multirow{5}{*}{Coding the experiment} & data preprocessing & Qingxia & 05/04 \\
    %& building probing system \& INLP & Saiya \& Qingxia & 05/25 \\
    %& building evaluation system & Gladys & 05/25 \\
    %& merging subparts together & Saiya & 06/02\\ 
    %& running different task pairs & Saiya & 06/02\\
    Original & 0 & p$\%$ & p$\%$ \\
    \hline
    %\multirow{2}{*}{Presenting related topic} & \multirow{2}{*}{presenting INLP method} & \multirow{2}{*}{Lindsay \& Gladys} & \multirow{2}{*}{05/11}\\
    %& & & \\
    Rand($|v_{nosem}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{nosyn}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{syn-sem}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{sem-syn}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{syn*sem}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{sem*syn}|$) & 0 & p$\%$ & p$\%$ \\
    \hline
    %\multirow{2}{*}{Writing paper} & \multirow{2}{*}{writing out experiment } & \multirow{2}{*}{Lindsay} & \multirow{2}{*}{06/09}  \\
    %& & & \\
    $v_{nosem}$ & 0 & p$\%$ & p$\%$ \\
    $v_{nosyn}$) & 0 & p$\%$ & p$\%$ \\
    $v_{syn-sem}$ & 0 & p$\%$ & p$\%$ \\
    $v_{sem-syn}$ & 0 & p$\%$ & p$\%$ \\
    $v_{syn*sem}$ & 0 & p$\%$ & p$\%$ \\
    $v_{sem*syn}$ & 0 & p$\%$ & p$\%$ \\
    \hline
    \hline
    \end{tabular}
    \caption{\label{role description} Description of how to read results
    }
    \end{table*}

>>>>

\citealp{disentangle} use paraphrase pairs and new target syntax to train a semantic encoder, syntactic encoder and decoder to learn separate representations of the semantic and syntactic information contained in BART embeddings, in order to create semantically equivalent paraphrases with the new syntactic structure. Alongside the encoders they also train an adversarial syntax discriminator to try and predict the source syntax from the semantic embeddings, thus encouraging the disentanglement of the semantic and syntactic information by training the semantic embedder to remove as much syntactic information as possible. Their results show that one can achieve some removal of syntactic information from semantic embeddings, so disentanglement of some information is possible. Though they do not achieve perfect separation of the two information sets. Other non-linear approaches to syntactic-semantic information disentanglement have been carried out in \citep{multiDis}

@@

%Need to update this section to reflect the new dataset
% might change later if choose to run experiment in silver
We use the English Parallel Meaning Bank v4.0 \citep{pmbData} to test the linear separability of the semantic and syntactic information in word-level BERT embeddings. This dataset is consist of gold standard and silver standard of word-level semantic tags. The gold standard contains 5,438 sentences with annotations that are manually verfied while the silver standard contains 62,739 sentences with autogenerated annotations. All of our experiments will be conducted on gold standard data.

The original dataset does not include CCG Tags, but \citealp{pmbData} utilized CCG parser to produce CCG tags. We follow a similar procedure to apply a CCG parser \citep{ccg_tagger} to develop word-level CCG tags. Once we obtain both CCG tags and semantics tags for the dataset, we can perform syntactic and semantics probing task as desired.%Fill this in
 %Do we drop anything or do any other pre-processing that I missed here?

@@

\begin{table*}[h]
    \centering
    \begin{tabular}{llll}
    \hline
    \textbf{Embedding} & \textbf{Directions Removed} & \textbf{CCG Tagging} &\textbf{Semantic Labeling} \\
    \hline
    Original & 0 & $84.75\%$ & $88.56\%$ \\
    \hline

    Rand($|v_{nosem}|$) & 77 & $84.06\%$ & $87.71\%$ \\
    Rand($|v_{nosyn}|$) & 159 & $82.26\%$ & $87.57\%$ \\
    Rand($|v_{syn-sem}|$) & 81 & p$83.11\%$ & $87.61\%$ \\
    Rand($|v_{sem-syn}|$) & 161 & $82.46\%$ & $86.04\%$ \\
    Rand($|v_{syn*sem}|$) & 0 & p$\%$ & p$\%$ \\
    Rand($|v_{sem*syn}|$) & 0 & p$\%$ & p$\%$ \\
    \hline
    $v_{nosem}$ & 77 & $27.93\%$ & $17.28\%$ \\
    $v_{nosyn}$ & 159 & $23.76\%$ & $49.43\%$ \\
    $v_{syn-sem}$ & 81 & $34.24\%$ & $33.91\%$ \\
    $v_{sem-syn}$ & 161 & $26.96\%$ & $50.21\%$ \\
    $v_{syn*sem}$ & 0 & p$\%$ & p$\%$ \\
    $v_{sem*syn}$ & 0 & p$\%$ & p$\%$ \\
    \hline
    \hline
    \end{tabular}
    \caption{\label{role description} Description of how to read results
    }
    \end{table*}
